---
title: "Analysis"
output: html_document
date: '2022-07-06'
---

# Setup 

```{r}
# IPUMS & Spatial Data Tools
library(tidyverse)
library(ipumsr) 
library(sf)
library(terra)
library(ggspatial)
```

```{r}
# Analysis Tools 
library(srvyr)
library(survey)
library(lme4)
library(broom)
library(broom.mixed)
```

# 1 - IPUMS PMA Data 

## Load IPUMS PMA data into R

```{r}
pma <- read_ipums_micro(
  ddi = "data/pma_00136.xml", # Use your own file name here 
  data = "data/pma_00136.dat.gz" # Use your own file name here 
)

pma
```

## Data format 

```{r}
pma %>% count(RESULTFQ_1, RESULTFQ_2)

pma <- pma %>% filter(RESULTFQ_1 == 1 & RESULTFQ_2 == 1)
```

# 2- GPS data 

## Load PMA GPS data into R 

```{r}
gps <- read_csv("data/pma_gps.csv")

gps
```

## GPS data as a Simple Features Object

```{r}
gps <- gps %>% 
  st_as_sf(
    coords = c("GPSLONG", "GPSLAT"), 
    crs = 4326
  ) 

gps
```

# 3 - Mapping Birth Outcomes

## Birth outcomes for individuals

```{r}
pma %>% count(PANELBIRTH_2)

pma %>% count(PANELBIRTH_2 == 1)
```

## Birth outcomes by EA 

```{r}
ea_summary <- pma %>% 
  as_survey_design(weight = PANELWEIGHT) %>% 
  group_by(ea = EAID_1, urban = URBAN == 1) %>% 
  summarise(birth_prop = survey_mean(PANELBIRTH_2 == 1, vartype = NULL)) %>% 
  ungroup() %>% 
  mutate(
    ntile = ntile(birth_prop, 2),
    many_births = ntile == 2
  )

ea_summary
```

## Merging GPS data 

```{r}
ea_summary_gps <- full_join(gps %>% select(ea = EAID), ea_summary, by = "ea")

ea_summary_gps
```

## Mapping births by EA

```{r, fig.align='center'}
ggplot() + 
  layer_spatial(ea_summary_gps, aes(shape = many_births, color = many_births)) + 
  theme_minimal()
```

## Adding a Shapefile

```{r, results='hide'}
shape <- st_read("data/geobf") %>% select(ADMIN_NAME)

shape
```

## Mapping births by EA

```{r, fig.align='center', `code-line-numbers` = "3"}
ggplot() + 
  layer_spatial(ea_summary_gps, aes(shape = many_births, color = many_births)) + 
  layer_spatial(shape, alpha = 0) + # `alpha` controls transparency 
  theme_minimal()
```

# 4 - CHIRPS: Annual rainfall summary 

## Load Raster Data into R

```{r}
june5_2020 <- rast("data/chirps/20200605.tif")

june5_2020
```

## June 5, 2020

```{r, `code-line-numbers`="5-7", fig.align='center'}
ggplot() + 
  layer_spatial(mask(june5_2020, vect(shape), touches = FALSE)) + 
  layer_spatial(shape, alpha = 0) + 
  theme_minimal() + 
  scale_fill_gradient2(
    low = "transparent", 
    high = "#4375B7", 
    na.value = "transparent"
  ) + 
  labs(fill = "Rainfall (mm)")
```

## Aggregation 

```{r, `code-line-numbers`="3-10"}
years <- map(1981:2020, ~{
  yr <- .x
  months <- map(6:9, ~{
      list.files(
        "data/chirps",
        full.names = TRUE,
        pattern = paste0(yr, "0", .x, ".*tif")
      ) 
    })
  unlist(months) 
})

years <- set_names(years, 1981:2020) 
```

## Raster Layers 

```{r}
years <- years %>% map(~.x %>% rast)

years$`2020`
```

## Seasonal Rainfall Accumulation

```{r}
years$`2020` %>% sum()

chirps_seasonal_sum <- map(years, ~.x %>% sum)
```

## One Layer per Year 

```{r}
chirps_seasonal_sum <- rast(chirps_seasonal_sum)

chirps_seasonal_sum
```

## How rainy was 2020? 

```{r}
chrips_avg <- mean(chirps_seasonal_sum)
chirps_sd <- stdev(chirps_seasonal_sum)
chirps_z <- (chirps_seasonal_sum - chrips_avg) / chirps_sd

chirps_z
```

## 2020 Seasonal Rainfall Accumulation 

```{r}
ggplot() + 
  layer_spatial(mask(chirps_z$`2020`, vect(shape), touches = FALSE)) + 
  layer_spatial(shape, alpha = 0) +
  theme_minimal() + 
  scale_fill_gradient2(
    low = "#D72B22", 
    high = "#4375B7", 
    na.value = "transparent"
  ) + 
  labs(fill = "Z-score (1981-2020)")
```

# 5 - Bringing it All Together 

## From Points to Buffers

```{r, echo = F} 
ea_summary_gps
```

## Projection to Meters 

```{r}
ea_summary_gps <- ea_summary_gps %>% st_transform(crs = 32630)
ea_summary_gps

shape <- shape %>% st_transform(crs = 32630)
shape
```

## Creating Buffers 

```{r, results='hide'}
ea_summary_buffer <- ea_summary_gps %>% st_buffer(if_else(.$urban, 2000, 5000))

ea_summary_buffer
```

## Intersecting Regional Boundaries

```{r}
test_ea <- 854131009 

ea_summary_buffer <- ea_summary_buffer %>% 
  st_intersection(shape) %>% 
  st_filter(ea_summary_gps)

ggplot() + 
  layer_spatial(ea_summary_buffer %>% filter(ea == test_ea), alpha = 0) + 
  layer_spatial(ea_summary_gps %>% filter(ea == test_ea), color = "red") + 
  annotation_spatial(shape, alpha = 0) + 
  theme_minimal()
```

## Back to arc-degrees

```{r}
ea_summary_buffer <- ea_summary_buffer %>% st_transform(crs = 4326)
shape <- shape %>% st_transform(crs = 4326)
ggplot() + 
  layer_spatial(shape, alpha = 0) + 
  layer_spatial(ea_summary_buffer, aes(fill = many_births), color = 0) + 
  theme_minimal()
```

## Spatially Weighted Averages

```{r}
test_buffer <- ea_summary_buffer %>% filter(ea == test_ea)

ggplot() + 
  layer_spatial(
    chirps_z$`2020` %>% 
      crop(test_buffer, snap = "out")
  ) + 
  layer_spatial(test_buffer, alpha = 0) + 
  theme_minimal() + 
  scale_fill_gradient2(
    high = "#4375B7", 
    limits = c(0, 2)
  )

test_z <- chirps_z$`2020` %>%
  extract(vect(test_buffer), weights = TRUE)

test_z

test_z %>% summarise(
  wtd_mean = weighted.mean(`2020`, weight)
)
```

## Enumeration Area Z-scores

```{r}
ea_summary_buffer <- ea_summary_buffer %>% rowid_to_column("ID")

ea_summary_z <- chirps_z$`2020` %>% 
  extract(vect(ea_summary_buffer), weights = TRUE) %>% 
  tibble() %>% 
  group_by(ID) %>% 
  summarise(z = weighted.mean(`2020`, weight)) 

ea_summary_z

ea_summary_z <- ea_summary_buffer %>% 
  st_drop_geometry() %>% 
  select(ID, ea) %>% 
  full_join(ea_summary_z, by = "ID")

ea_summary_z
```

# 6 - Analysis

## Join Z-scores with Original PMA Data 

```{r}
pma_chirps <- pma %>% 
  transmute(
    ea = EAID_1, 
    wt = PANELWEIGHT,
    birth = PANELBIRTH_2 == 1,
    urban = URBAN == 1
  ) %>% 
  left_join(ea_summary_z, by = "ea") 

pma_chirps
```

## Binomial model with cluster-robust SEs

```{r}
# Weighted with cluster-robust SEs
pma_chirps %>%
  as_survey_design(weight = wt, id = ea) %>%
  svyglm(birth ~ z*urban, design = ., family = "quasibinomial") %>%
  tidy(exp = TRUE, conf.int = TRUE) %>%
  mutate(sig = gtools::stars.pval(p.value))
```

## Multilevel model with EA random effects 

```{r}
glmer(
  birth ~ z*urban + (1 | ea), 
  data = pma_chirps,
  family = "binomial") %>%
  tidy(exp = TRUE, conf.int = TRUE) %>%
  mutate(sig = gtools::stars.pval(p.value))
```

